Let There Be Light: \\ Early Observations of the Normal Type Ia Supernova \abc\ \\Show No Evidence for a Dark Period

### 2017-06-30 doc/ms.tex
Add more stringent constraints on the companion interaction scenario to the paper.

The detection of excess emission due to the collision of SN ejecta 
with a non-degenerate companion requires a favorable orbital 
alignment relative to the line of sight. Thus, from geometric 
considerations alone the probability of detecting ejecta-companion 
interaction is low, $\sim$10\%. For any single type Ia SN, this 
probability is further reduced because many explosions come from 
WD-WD binaries. \citet{2010ApJ...708.1025K} calculates that the 
collision of SN ejecta with a companion  
generates thermal emission with a spectrum that peaks in the 
UV. Thus, any resulting \textit{g}-band emission is very weak.

To examine the possibility of a SN-companion signature in the early
light curve of \abc, we employ the \citet{2010ApJ...708.1025K}
model and assume canonical ejecta mass of $1.4\sm$, expansion velocity of
$10^{4}\,\textrm{km}\,\textrm{s}^{-1}$, and constant opacity of
$0.2\,\textrm{cm}^2\,\textrm{g}^{-1}$. We calculate the expected 
$g_\mathrm{PTF}$ brightness of an ejecta-companion collision at the 
distance of \abc\ using the parameterized equations in 
\citet{2012ApJ...749...18B}. If we assume the binary is aligned with 
the optimal orientation relative to the line of sight, the binary 
separation would need to exceed $2\times10^{14}\,\textrm{cm}$ to 
explain the initial detection of \abc. Assuming the companion fills 
its Roche lobe, it would need to 
have a radius of $\sim10^{14}\,\textrm{cm}\simeq10 \,\mathrm{AU}$.
Such a radius is implausibly large for a WD binary companion, 
therefore, we conclude that the early emission from \abc\ is not 
due to a collision between the ejecta and a non-degenerate companion. Our early \textit{Swift} observations support this conclusion, as the UV $-$ optical colors are very red, whereas iPTF\,14atg exhibited much bluer UV $-$ optical colors at the same epoch \citep{2015Natur.521..328C}.

% \begin{figure}[!thb]
%   \centering
%   \includegraphics[width=0.45\textwidth]{SNCompanion.pdf}
%   \caption{Apparent magnitude of a SN-companion collision (gray region)
%     against the first detection of \abc\ (black). \amiller{Cut this
%     figure? Doesn't add a lot of information.}}
%   \label{fig:SN-companion}
% \end{figure}

>>>>

The detection of excess emission due to the collision of the SN ejecta 
with a non-degenerate companion requires a favorable orbital 
alignment relative to the line of sight. Thus, from geometric 
considerations alone the probability of detecting ejecta-companion 
interaction is low, $\sim$10\%. \citet{2010ApJ...708.1025K} calculates that the 
collision of SN ejecta with a companion  
generates thermal emission with a spectrum that peaks in the 
UV. The resulting \textit{g}-band emission is expected to be weak.

To examine the possibility of a SN-companion signature in the early
light curve of \abc, we employ the \citet{2010ApJ...708.1025K}
model and assume canonical values for the ejecta mass, $1.4\,\sm$, expansion velocity,
$10^{4}\,\textrm{km}\,\textrm{s}^{-1}$, and a constant opacity, 
$0.2\,\textrm{cm}^2\,\textrm{g}^{-1}$. We calculate the expected 
$g_\mathrm{PTF}$ brightness of an ejecta-companion collision at the 
distance of \abc\ behind a total redenning of $E(B-V) = 0.0779 \, \mathrm{mag}$ using the parameterized equations in 
\citet{2012ApJ...749...18B}. If we assume the binary is aligned with the optimal orientation relative to the line of sight, a binary separation of $a \approx 2 \times 10^{13}\, \mathrm{cm}$ is needed to explain the initial detection of \abc, as shown in Figure~\ref{fig:SN-companion}. This model does not, however, match the $g_\mathrm{PTF}$ evolution for $t > t_0 + 1 \, \mathrm{d}$ (though it is possible that the SN photosphere dominates the companion-interaction signature at this phase). Furthermore, this model significantly overpredicts the observed flux in the \textit{Swift}/$UVW1$ filter $\sim$3.1~d after $t_0$. As such, we do not favor such a model as the explanation for the early emission from \abc.

Figure~\ref{fig:SN-companion} additionally shows that a companion at $a = 10^{14} \, \mathrm{cm}$ provides a good match to the observations $\sim$1--3~d after $t_0$ if $t_\mathrm{exp} \approx t_0 + 0.3 \; \mathrm{d}$. Like the $a = 2 \times 10^{13} \, \mathrm{cm}$ model, this model significantly overpredicts the observed UV flux from \abc. Additionally, this model cannot account for the brightness of \abc\ at the epoch of discovery. The challenges associated with each of the previous two models lead us to conclude that the early evolution of \abc\ cannot be explained via the interaction of the SN ejecta with a red-giant companion.\footnote{The \citet{2010ApJ...708.1025K} model assumes Roche-lobe overflow, meaning that $a > 10^{13} \, \mathrm{cm}$ implies a companion with $R_\ast \gtrsim 100 \, R_\odot$.} We cannot, however, exclude the presence of a red-giant companion as our calculations have assumed that the binary is aligned with the optimal geometry relative to the line of sight. If the geometry is not favorable, then it is possible that signatures from interaction with a companion are not visible.

 % Our early \textit{Swift} observations support this conclusion, as the UV $-$ optical colors are very red, whereas iPTF\,14atg exhibited much bluer UV $-$ optical colors at the same epoch \citep{2015Natur.521..328C}.

\begin{figure}[!thb]
  \centering
  \includegraphics[width=0.48\textwidth]{companion_lc.pdf}
  \caption{Comparison of SN ejecta-companion interaction models 
  with the early $g_\mathrm{PTF}$ light curve of \abc. 
  Detections are shown as circles, while 3$\sigma$
  upper limits are shown as downward facing arrows.
  expected $g_\mathrm{PTF}$ light curve for the smallest orbital 
  separation capable of explaining the initial detection of \abc. 
  This model also requires $t_\mathrm{exp} \approx t_0 - 1 \; 
  \mathrm{d}$. Assuming $t_\mathrm{exp} = t_0$, interaction with a 
  companion at $a = 2 \times 10^{13} \; \mathrm{cm}$ matches the 
  initial detection of \abc\ (solid line). This model 
  significantly overestimates the UV flux detected from \abc\ 
  $\sim$3 d after explosion (see text). The $g_\mathrm{PTF}$ 
  observations $\sim$1--3~d after $t_0$ are well matched to 
  companion-interaction models for $a = 10^{14} \; \mathrm{cm}$ 
  if $t_\mathrm{exp} \approx t_0 + 0.3 \; \mathrm{d}$ (dash-dot 
  line). This model significantly underestimates the flux at the 
  epoch of discovery, while also overestimating the UV flux. Thus, 
  none of these possibilities uniquely explain the early flux from 
  \abc.
  \amiller{Add Swift light curve and photometry to make it clear that two large companions don't explain the observations? Also -- make labels for this fig bigger}}
  \label{fig:SN-companion}
\end{figure}

Finally, we examine the minimum binary separation capable of explaining the observed brightness at the epoch of discovery. Figure~\ref{fig:SN-companion} shows that models with $a \approx 8 \times 10^{11} \, \mathrm{cm}$ peak at $\sim$21.5~mag in the $g_\mathrm{PTF}$-band, provided that $t_\mathrm{exp} \approx t_0 - 1 \; \mathrm{d}$. To explain just the epoch of discovery, models with larger $a$ would require even earlier explosion times, which is not compatible with our non-detections the night before discovery, while models with smaller $a$ are not bright enough. While it is otherwise compatible with the observations, we do not favor the $a \approx 8 \times 10^{11} \, \mathrm{cm}$ model as the explanation for the early flux from \abc\ because it does not explain the 2 mag rise in the $\sim$24 hr after discovery.

Recently, \citet{2017arXiv170608990H} showed that the early light curve of SN\,2017cbv has a clearly resolved bump in the $U$, $B$, and $g'$ bands in the $\sim$5 d after explosion. \citeauthor{2017arXiv170608990H} find that the bump can be explained via the combination of ejecta-companion interaction and the normal evolution of a type Ia SN. A challenge for this model is that it significantly overpredicts the UV brightness of the SN compared to what is observed. It is argued in \citet{2017arXiv170608990H} that the shock resulting from the companion collision may not be a blackbody, which could reconcile the model with the observed UV photometry. Qualitatively, \abc\ and SN\,2017cbv exhibit similar characteristics: both feature strong \ion{C}{2} absorption in their early spectra and a bump in their early light curves. Below, we argue for circumstellar interaction and strong $^{56}$Ni mixing as a possible explanation for \abc, and indeed \citet{2017arXiv170608990H} consider these possibilities for SN\,2017cbv as well. While there are arguments in favor of and against each of these possibilities, we believe that the early behavior of SN\,2017cbv and \abc\ reflect a common physical origin.
\amiller{should this paragraph be in the conclusions?}
